---
title: "IMDb_project"
author: "Mario Catuogno & Francesco Pastore"
date: "27 September 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Project Overview

The program uses a set of parameters, to cluster and classify movies and propose one that you haven't seen yet. It is possible to catalogue previous seen movies to help the program to better select the results.

The project have a long-term goal, we have just started the descriptive analysis of the huge dataset of **IMDb** (*more than 12Gb of compressed data*) and we are still in the **Data Cleansing** stage.

## Project Roadmap

Goal | Description | Status
--- | --- | ---
Import | Import the following dataset: **actors.csv**, **actresses.csv**, **directors.csv**, **editors.csv**, **genres.csv**, **movies.csv**, **plots.csv**, **running-time.csv** | 30%

## R Functions list

* get_history
    * intial_history
    * suggestion_history
* get_mood_social_context
* get_date
* get_trend
    * imdb_updates
* sentiment_complete_analysis
    * sentiment_cluster
    
## Functions description

The function `get_history()` allows to retrieve 
```{r }

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
